\documentclass{article}
%\documentclass[prl,amsmath,amssymb]{revtex4} % PRL

% margins of 1 inch:
\setlength{\topmargin}{-.5in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{0in}
\setlength{\textwidth}{6.5in}

\usepackage[pdftex]{hyperref} % hyperlink equation and bibliographic citations
\usepackage[dvips]{graphicx,color}
\usepackage{amsmath} % advanced math
\usepackage{verbatim}
\usepackage{natbib} % bibilography 
\usepackage{mciteplus} % collapse multiple citations in bibilography
\usepackage{multicol}
\usepackage[toc,page]{appendix} % http://tex.stackexchange.com/questions/49643/making-appendix-for-thesis


% from http://www.flakery.org/search/show/569
%\newcommand{\infint}{\ensuremath{\int_{-\infty}^{\infty}}}
\newcommand{\cf}{\textit{c.f.}} % "compare". In context the abbreviation advises readers to consult other material, drawing attention to related ideas that provide additional arguments or information.
\newcommand{\ie}{\textit{i.e.}} % i.e. is used to explain, clarify or rephrase a statement
\newcommand{\eg}{\textit{e.g.}} % “for the sake of example”. Used to introduce an example or list of examples to illustrate what is being discussed.
\newcommand{\eqn}[1]{Eq.\ (\ref{#1})}
\newcommand{\pfrac}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\pdg}{Physics Derivation Graph}

\begin{document}
\title{User Manual for the Physics Derivation Graph}

\author{Ben Payne$^{1}$\footnote{Corresponding author: ben.is.located@gmail.com}, Michael Goff$^{2}$\\
{\it $^{1}$Department of Fun, University Name \& Town, city, State Zip}\\
{\it $^{2}$Department of Mathematics, University of Maryland \& Baltimore 21228}}

\date{\today}


\maketitle % declares end of title page
\begin{abstract}
Overview of current implementation. It is assumed here that the Introduction to the \pdg has been read. 
\end{abstract}

\begin{multicols}{2}

%\tableofcontents

%\newpage

\section{Introduction}

% relevant project background
The \pdg is a project designed to capture mathematical physics knowledge. The mathematics necessary to describe physics can be written down, in this case in the form of a graph. Thus, the core of this project is a set of databases storing that content. There are a set of scripts for various manipulations of the content. Those scripts include the ability to generate visualizations of the graph, and to generate reports based on the graph content.

 Section~\ref{sec:databases} describes the databases which store these values. 

\section{Databases\label{sec:databases}}

The content associated with physics knowledge takes the form of expressions (\textit{equations and inequalities}), symbols, and inference rules. \textit{An inference rule is an atomic transformation of one expression to another.} An example is provided here. 

Frequency $f$ and period $T$ are related by
\begin{equation}
T\ f = 1
\label{eq:period_and_freq}
\end{equation}
Thus, frequency in terms of the period is
\begin{equation}
f = 1/T
\label{eq:freq_is_inverse_period}
\end{equation}
The relation between \eqn{eq:period_and_freq} and \eqn{eq:freq_is_inverse_period} is that both sides of \eqn{eq:period_and_freq} were divided by $T$. 

In this example, there are two mathematical expressions: \eqn{eq:period_and_freq} and \eqn{eq:freq_is_inverse_period}. These expressions are related by an inference rule: ``Divide both sides of first equation by a value to yield the second equation.'' This inference rule takes an argument, referred to here as the ``feed'', which in this example is $T$. This set of steps is shown graphically in~Fig.~\ref{fig:freq_period}. This simple step between two expressions is documented in five separate databases: expressions, symbols, feeds, inference rules, and connections. 

Each expression, symbol, and inference rule appears only once in the respective database. Each time an expression, symbol, or inference rule is used in a derivation, that unique instance is referenced. This referencing of unique expressions, symbols, and inference rules is done using a numeric identifier (alphanumeric for inference rules).

In the expressions database, each expression has an associated 10 digit number in base 10. For example, \eqn{eq:freq_is_inverse_period} is 2113211456 and \eqn{eq:period_and_freq} is 2131616531. Similarly, the symbols have unique 15 digit numeric identfiers. Period $T$ is 192938440120938 and frequency $f$ is 102938475990112. The database of expressions references these symbol identifiers -- expression 2113211456 uses symbols 192938440120938 and 102938475990112. 

The connections database is the set of derivations. A derivation involving \eqn{eq:freq_is_inverse_period} refers to the expression identifier 2113211456 and relates it to 2131616531 by an inference rule which is identified by the alphanumeric string ``dividebothsidesby''. This inference rule takes an argument, referred to here as a ``feed.'' The feed database contains the value, in this example $T$ and a 7 digit identifier 8837284. Feeds are not unique, though the identifiers used are. The $T$ referred to for this use of the inference rule ``dividebothsidesby'' does not need to be associated with other instances of ``dividebothsidesby''. 

The \pdg is designed to show one instance of each expression, but feeds and inference rules will have duplicates on the graph. To allow for this, the connections database has a unique 7 digit numeric identifier assocaited with each use of an inference rule. 

\subsection{Implementation\label{sec:implementation}}
The databases supporting the \pdg are implemented using XML\cite{2008_XML}. 

\section{Database manipulation}

The databases are manipulated using software written in Python\cite{2015_Python} version 2.7.6. These scripts are described here in order of increasing complexity. 

\subsection{Listing}

The two simplest scripts are for listing content from the connections and inference rules databases, respectively. 
\begin{verbatim}
bin/list_connection_sets.py
bin/list_inference_rules.py
\end{verbatim}

\subsection{Statistical Analysis}
There are three scripts which build on the listing functionality and carry out simple analysis of the counts. 
\begin{verbatim}
popularity_of_inference_rules.py
popularity_of_statements.py
popularity_of_symbols.py
\end{verbatim}

\begin{verbatim}
bin/popularity_of_statements.py
\end{verbatim}
For example, \eqn{eq:freq_is_inverse_period} is referenced twice as of this writting. 

\subsection{Utilities}

\begin{verbatim}
create_pdf_of_statements.py
create_pictures_of_inference_rules.py
create_pictures_of_statements_and_feeds.py
generate_new_random_index.py
\end{verbatim}

\subsection{Primary functions}

\begin{verbatim}
check_connections_using_CAS_sympy.py
build_connection_set_pdf.py
build_connections_graph.py
\end{verbatim}

The \texttt{build\_*.py} scripts are described in detail in sections \ref{sec:reports} and \ref{sec:visualization}, respectively. 

\section{Report generation\label{sec:reports}}

\LaTeX\cite{1994_Latex} PDFs

\section{Visualization of Graph\label{sec:visualization}}

DOT language\cite{2015_Graphviz} 

\section{Summary}

\section{Bibliography}

\bibliographystyle{unsrt}
\bibliography{../bibliography} % external bibtex flat-file database
\end{multicols}

\newpage
\appendix
%\begin{appendices}
%\include{appendix_test_cases}
%\end{appendices}


\end{document}
